### 2014-04-02 Docs/flowchart/flowchart.tex
start of an index just for runtime parameters
more runtime parameters to the index

  There are a variety of choices of quantities to predict to the
  edges (controlled by {\tt species\_pred\_type}---see Chapter~\ref{ch:pert}).
  By default, we use the equations 
\begin{eqnarray}
\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
     \label{eq:Perturbational Density}  \\
\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
     \omegadot_k. \label{eq:Primitive Species}
\end{eqnarray}
  to
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density to full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  There are a variety
   of quantities that we can predict to the interfaces here (controlled
   by {\tt enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
   the default case, we use the perturbational
  enthalpy equation, neglecting reactions,
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
    \label{eq:Perturbational Enthalpy}.
\end{equation}
  to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.

@@

This change comes into \maestro\ in a variety of steps, summarized here.

@@

%-----------------------------------------------------------------------------
% graphical flowcharts
%-----------------------------------------------------------------------------
\begin{figure}[tb]
\centering
\includegraphics[scale=0.65]{\flowfigpath/flowchart}
\caption[Graphical flowchart of \maestro]
  {\label{Fig:flowchart} A flowchart of the algorithm.  The
  thermodynamic state variables, base state variables, and local velocity are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  The predictor-corrector steps are
  outlined by the dotted box.  The blue text indicates state
  variables that are the same in {\bf Step 6} as they are in
  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
  The diffusion steps (4a and 8a) are optional, depending on
  {\tt use\_thermal\_diffusion}.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tb]                                               
\centering
\includegraphics[scale=0.75]{\flowfigpath/flowchart_4_8}
\caption[Graphical flowchart of the density and enthalpy update steps]
{\label{Fig:flowchart48} A flowchart for {\bf Steps 4} and {\bf 8}.
  The thermodynamic state variables and base state variables are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  Note, for {\bf Step 4}, the updated
  quantities should also have a $\star$ superscript, e.g., {\bf Step
    8I} defines $T^{(2)}$ while {\bf Step 4I} defines $T^{(2),\star}$
  .}
\end{figure}

@@

\item {\em Initial Projection}: if {\tt do\_initial\_projection = T}, then we 
   first project the velocity field with $\rho = 1$ and $\beta_0^0$.
   The initial projection does not see reactions
   or external heating, and thus we set $\dot\omega = \Hnuc = \Hext = 0$ in $S$.
   The reason for ignoring reactions and heating is that we need some kind of 
   time scale over which to compute the effect of reactions, but we first 
   need an estimate of the velocity
   field in order to derive the time step that will be used as a time scale.
   The elliptic equation we solve is 
   \begin{equation}
   \nabla \cdot \beta_0^0 \nabla \phi = \underbrace{\beta_0^0(S - \Sbar)}_\mathrm{0~ except~ for~ diffusion} - \nabla \cdot (\beta_0^0 \Ub^{\initp})
   \end{equation}
   This $\phi$ is then used to correct the velocity field to obtain $\Ub^{0,0}$.
   If {\tt do\_initial\_projection = F}, set $\Ub^{0,0} = \Ub^{\initp}$.

\item {\em ``Divu'' iterations}: Next we do {\tt init\_divu\_iter} iterations 
  to project the velocity field using a constraint that sees reactions
  and external heating. 
  The initial timestep estimate is provided by {\tt firstdt} and
  $\Ub^{0,0}$, to allow us to compute the effect of reactions over $\Delta t/2$.

@@

Next, we need to construct approximations to $\etarho^{-\myhalf}, \psi^{-\myhalf}, S^1$,
and $\pi^{-\myhalf}$.  As initial approximations, set 
$\etarho^{-\myhalf}=0, \psi^{-\myhalf}=0, S^{1,0}=S^0$, and $\pi^{-\myhalf}=0.$
\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
\renewcommand{\labelenumii}{\roman{enumii}.}
\addtocounter{enumi}{2}
\item {\it Pressure iterations}: Here we do {\tt init\_iter} iterations to get an
  approximation for the lagged pressure:

@@

\section{Changes from Earlier Implementations}
\subsection{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item We defined the mapping of data between a 1D radial array and the 3D Cartesian
grid for spherical problems (which we improve upon in the multilevel paper).
\item We update $T$ after the call to {\bf React State}.
\item We have created a {\tt burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to {\tt base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We have an option, controlled by {\tt use\_tfromp}, to update temperature 
using $T=T(\rho,X_k,p_0)$ rather than $T=T(\rho,h,X_k)$.  The former completely 
decouples enthalpy from our system.  For spherical problems, we use 
{\tt use\_tfromp = TRUE}, for planar problems, we use {\tt use\_tfromp = FALSE}.
\item For spherical problems, we have changed the discretization of 
$\Ubt\cdot\nabla p_0$ in the enthalpy update to 
$\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
\item In paper III we discretized the enthalpy evolution equation in
terms of $T$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use {\tt enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using {\tt do\_eos\_h\_above\_cutoff = TRUE}.
\end{enumerate}

@@

\subsection{Changes Between Paper 5 and the XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by {\tt use\_thermal\_diffusion},
{\tt temp\_diffusion\_formulation}, and {\tt thermal\_diffusion\_type}.
\item We added the volume discrepancy term to the velocity constraint equation,
controlled by the input parameter, {\tt dpdt\_factor}.
\item For certain problems, we need to set {\tt do\_eos\_h\_above\_cutoff = FALSE}
to prevent large, unphysical velocities from appearing near the edge of the star.
\end{enumerate}

@@

\begin{itemize}
\item Should we use a predictor-corrector for updating the full-state density?
Specifically, after calling {\bf Correct Base}, should we do a full-state density 
advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
\item We are still exploring the effects of {\tt use\_tfromp = FALSE} for spherical
problems.  We would eventually like to run in this mode, but $T=T(\rho,X_k,p_0)$ 
and $T=T(\rho,h,X_k)$ drift away from each other more than we would like.  Our
attempts at incorporating a {\tt dpdt\_factor} for spherical problems have not 
been successful.
\end{itemize}

>>>>

  There are a variety of choices of quantities to predict to the
  edges (controlled by \runparam{species\_pred\_type}---see Chapter~\ref{ch:pert}).
  By default, we use the equations 
\begin{eqnarray}
\frac{\partial\rho'}{\partial t} &=& -\Ub\cdot\nabla\rho' - 
     \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right),
     \label{eq:Perturbational Density}  \\
\frac{\partial X_k}{\partial t} &=& -\Ub\cdot\nabla X_k + 
     \omegadot_k. \label{eq:Primitive Species}
\end{eqnarray}
  to
  predict $\rho^{'(1)} = \rho^{(1)} - \rho_0^n$ and 
  $X_k^{(1)} = (\rho  X_k)^{(1)} / \rho^{(1)}$ to time-centered edges using 
  $\Ub = \uadvone+w_0^{\nph,\star}\eb_r$, yielding $\rho^{'\nph,\star,\pred}$ 
  and $X_k^{\nph,\star,\pred}$.
  We convert the perturbational density to full state density using
\begin{equation}
\rho^{\nph,\star,\pred} = \rho^{'\nph,\star,\pred} + \frac{\rho_0^n + \rho_0^{(2a),\star}}{2},
\end{equation}
  where the base state density terms are mapped to Cartesian edges.
  Then,\\
  $(\rho X_k)^{\nph,\star,\pred} = (\rho^{\nph,\star,\pred}X_k^{\nph,\star,\pred})$.

@@

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\star,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.  There are a variety
   of quantities that we can predict to the interfaces here (controlled
   by \runparam{enthalpy\_pred\_type}---see Chapter~\ref{ch:pert}).  For
   the default case, we use the perturbational
  enthalpy equation, neglecting reactions,
\begin{equation}
\frac{\partial(\rho h)'}{\partial t} = -\Ub\cdot\nabla(\rho h)' - 
   (\rho h)'\nabla\cdot\Ub - \nabla\cdot\left[(\rho h)_0\Ubt\right] + \Ubt\cdot\nabla p_0 
    \label{eq:Perturbational Enthalpy}.
\end{equation}
  to predict
  $(\rho h)' = (\rho h)^{(1)} - (\rho h)_0^n$ to time-centered edges, 
  using $\Ub = \uadvone+w_0^{\nph,\star} \eb_r$,
  yielding $(\rho h)^{'\nph,\star,\pred}$.  We convert the perturbational 
  enthalpy to a full state enthalpy using
\begin{equation}
(\rho h)^{\nph,\star,\pred} = (\rho h)^{'\nph,\star,\pred} + \frac{(\rho h)_0^n + (\rho h)_0^{n+1,\star}}{2}.
\end{equation}
  For planar geometry, we map $(\rho h)_0$ directly to Cartesian edges.
  In spherical geometry, our experience has shown that a slightly different
  approach leads to reduced discretization errors.  We first map 
  $h_0 \equiv (\rho h)_0/\rho_0$ and $\rho_0$ to Cartesian edges separately, 
  and then multiply these terms to get $(\rho h)_0$.

@@

This change comes into \maestro\ in a variety of steps, summarized here.
To enable this portion of the algorithm, set \runparam{use\_delta\_gamma1\_term} {\tt = T}.

@@

%-----------------------------------------------------------------------------
% graphical flowcharts
%-----------------------------------------------------------------------------
\begin{figure}[tb]
\centering
\includegraphics[scale=0.65]{\flowfigpath/flowchart}
\caption[Graphical flowchart of \maestro]
  {\label{Fig:flowchart} A flowchart of the algorithm.  The
  thermodynamic state variables, base state variables, and local velocity are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  The predictor-corrector steps are
  outlined by the dotted box.  The blue text indicates state
  variables that are the same in {\bf Step 6} as they are in
  {\bf Step 2}, i.e., they are unchanged by the predictor steps.
  The diffusion steps (4a and 8a) are optional, depending on
  \runparam{use\_thermal\_diffusion}.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tb]                                               
\centering
\includegraphics[scale=0.75]{\flowfigpath/flowchart_4_8}
\caption[Graphical flowchart of the density and enthalpy update steps]
{\label{Fig:flowchart48} A flowchart for {\bf Steps 4} and {\bf 8}.
  The thermodynamic state variables and base state variables are
  indicated in each step.  Red text indicates that quantity was
  updated during that step.  Note, for {\bf Step 4}, the updated
  quantities should also have a $\star$ superscript, e.g., {\bf Step
    8I} defines $T^{(2)}$ while {\bf Step 4I} defines $T^{(2),\star}$
  .}
\end{figure}

@@

\item {\em Initial Projection}: if \runparam{do\_initial\_projection} {\tt = T}, then we 
   first project the velocity field with $\rho = 1$ and $\beta_0^0$.
   The initial projection does not see reactions
   or external heating, and thus we set $\dot\omega = \Hnuc = \Hext = 0$ in $S$.
   The reason for ignoring reactions and heating is that we need some kind of 
   time scale over which to compute the effect of reactions, but we first 
   need an estimate of the velocity
   field in order to derive the time step that will be used as a time scale.
   The elliptic equation we solve is 
   \begin{equation}
   \nabla \cdot \beta_0^0 \nabla \phi = \underbrace{\beta_0^0(S - \Sbar)}_\mathrm{0~ except~ for~ diffusion} - \nabla \cdot (\beta_0^0 \Ub^{\initp})
   \end{equation}
   This $\phi$ is then used to correct the velocity field to obtain $\Ub^{0,0}$.
   If {\tt do\_initial\_projection = F}, set $\Ub^{0,0} = \Ub^{\initp}$.

\item {\em ``Divu'' iterations}: Next we do \runparam{init\_divu\_iter} iterations 
  to project the velocity field using a constraint that sees reactions
  and external heating. 
  The initial timestep estimate is provided by {\tt firstdt} and
  $\Ub^{0,0}$, to allow us to compute the effect of reactions over $\Delta t/2$.

@@

Next, we need to construct approximations to $\etarho^{-\myhalf}, \psi^{-\myhalf}, S^1$,
and $\pi^{-\myhalf}$.  As initial approximations, set 
$\etarho^{-\myhalf}=0, \psi^{-\myhalf}=0, S^{1,0}=S^0$, and $\pi^{-\myhalf}=0.$
\begin{enumerate}
\renewcommand{\theenumi}{{\bf \alph{enumi}}}
\renewcommand{\labelenumii}{\roman{enumii}.}
\addtocounter{enumi}{2}
\item {\it Pressure iterations}: Here we do \runparam{init\_iter} iterations to get an
  approximation for the lagged pressure:

@@

\section{Changes from Earlier Implementations}
\subsection{Changes Between Paper 3 and Paper 4}
\begin{enumerate}
\item We defined the mapping of data between a 1D radial array and the 3D Cartesian
grid for spherical problems (which we improve upon in the multilevel paper).
\item We update $T$ after the call to {\bf React State}.
\item We have created a \runparam{burning\_cutoff\_density}, where the burning does
not happen below this density.  It is presently set to \runparam{base\_cutoff\_density}.
\item Use corner coupling in advection.
\item We have an option, controlled by \runparam{use\_tfromp}, to update temperature 
using $T=T(\rho,X_k,p_0)$ rather than $T=T(\rho,h,X_k)$.  The former completely 
decouples enthalpy from our system.  For spherical problems, we use 
{\tt use\_tfromp = TRUE}, for planar problems, we use {\tt use\_tfromp = FALSE}.
\item For spherical problems, we have changed the discretization of 
$\Ubt\cdot\nabla p_0$ in the enthalpy update to 
$\nabla\cdot(\Ubt p_0) - p_0\nabla\cdot\Ubt$.
\item In paper III we discretized the enthalpy evolution equation in
terms of $T$.  Since then we have discovered that 
discretizing the enthalpy evolution in perturbational form, $(\rho h)'$,
leads to better numerical properties.  We use \runparam{enthalpy\_pred\_type = 1}.
This is more like paper II.
\item We have turned off the evolution of $h$ above the atmosphere and instead
compute $h$ with the EOS using \runparam{do\_eos\_h\_above\_cutoff} {\tt = T}.
\end{enumerate}

@@

\subsection{Changes Between Paper 5 and the XRB Paper}
\begin{enumerate}
\item We have added thermal diffusion, controlled by \runparam{use\_thermal\_diffusion},
\runparam{temp\_diffusion\_formulation}, and \runparam{thermal\_diffusion\_type}.
\item We added the volume discrepancy term to the velocity constraint equation,
controlled by the input parameter, \runparam{dpdt\_factor}.
\item For certain problems, we need to set \runparam{do\_eos\_h\_above\_cutoff} {\tt = F}
to prevent large, unphysical velocities from appearing near the edge of the star.
\end{enumerate}

@@

\begin{itemize}
\item Should we use a predictor-corrector for updating the full-state density?
Specifically, after calling {\bf Correct Base}, should we do a full-state density 
advance and {\bf Correct Base} using the more accurate estimate of $\rho_0^{n+1}$?
\item We are still exploring the effects of \runparam{use\_tfromp}  {\tt = F} for spherical
problems.  We would eventually like to run in this mode, but $T=T(\rho,X_k,p_0)$ 
and $T=T(\rho,h,X_k)$ drift away from each other more than we would like.  Our
attempts at incorporating a \runparam{dpdt\_factor} for spherical problems have not 
been successful.
\end{itemize}
